$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2014-03-05 JournalToSubmit2013/modelingWithPISODM.tex
undergoing version of section 3 ... 05.04.2014 at 15:44

The \textit{$\pi$-UseCase} meta-model extends the UML Use Case meta-model for describing functional and non-functional requirements. 
Non-functional requirements are defined as \textit{constraints} over actions and data. The \textit{$\pi$-UseCase} meta-model describes the functional aspects of the application being developed and the restrictions that applies to each of them. These restrictions are modeled using the concepts: (i) constraint type and (ii) non-functional requirement with its specific attribute.
All constraints are described by stereotyped use cases. Every relationship between an use
case and a constraint is done through a dependency relationship (from UML).  A constraint is defined by a stereotype, containing its type and name. Constraints are represented by a stereotyped use case (´constraintª).
Each use case is related with one or more system requirements. Moreover, these use cases are related to non-functional requirements necessary for the proper execution of the application function. The function described by a use case can be part of a business service (in the case of service remote calls). This information is described through comments.

@@

 We use part of our case study to illustrate this meta-model. Each use case may have multiple constraints. In this example the publish music use case has an authentication constraint (user must authenticate). For a song to be published on any social network it is necessary to provide data privacy. The constraint type concerns the authentication values and the constraint has an non-functional attributes. The use case is also related with requirement, a business service and a non-functional requirement properties.
%{\color{red} Talk about this model for the example.}
\end{example}

@@

The \textit{$\pi$-ServiceProcess} meta-model defines the concept of \textit{service contract} to represent constraints over data and actions 
gathers the constraints described in the \textit{$\pi$-UseCase}.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusic3}
{\color{red} Talk about this model for the example.}
\end{example}

>>>>

The \textit{$\pi$-UseCase} meta-model extends the UML Use Case meta-model for describing functional and non-functional requirements. 
Non-functional requirements are defined as \textit{constraints} over actions and data. The \textit{$\pi$-UseCase} meta-model describes the functional aspects of the application being developed and the restrictions that applies to each of them. These restrictions are modeled using the concepts: (i) constraint type and (ii) non-functional requirement with its specific attribute.
All constraints are described by stereotyped use cases. Every relationship between an use
case and a constraint is done through a dependency relationship (from UML).  A constraint is defined by a stereotype, containing its type and name. Constraints are represented by a stereotyped use case («constraint»).
Each use case is related with one or more system requirements. Moreover, these use cases are related to non-functional requirements necessary for the proper execution of the application function. The function described by a use case can be part of a business service (in the case of service remote calls). This information is described through comments.

@@

 We use part of our case study to illustrate this meta-model. The ?To Publish Music? use case can update the Face- book or Twitter user status. Therefore, it is necessary to perform a social network authentication with the user?s data. Each social network uses different services and different forms of authentication. The authentication constraint is required to update a music status. The restriction is stereotyped as a «value» constraint, because the user?s id and password are verified.  Figure \ref{fig:CIM:tpmbpmn}  the buy music, download music, listen music and pay use cases. The process of buying a song requires the user?s private data for a Spotify account, and also a secure connection, represented as «value» and «business» constraint, respectively. For payment, the user must provide the data of payment card or PayPal account login and password, represented as «value» stereotype . Other restriction is that the minimum payment value is 2 euros.
%{\color{red} Talk about this model for the example.}
\end{example}

@@

The \textit{$\pi$-ServiceProcess} meta-model defines the concept of \textit{service contract} to represent constraints over data and actions 
gathers the constraints described in the \textit{$\pi$-UseCase}. A \textit{$\pi$-ServiceProcess} model represents the service processes and its contracts. It models the set of system activities and describe contracts for each activity or the process as a whole. Each service identified in the previous model (\textit{$\pi$-UseCase}) is mapped to an action. This model describes the workflow for the modelled application behaviour. The behaviour is represented by each action and the restrictions by the assertion description. All assertions are described by stereotyped actions.

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusic3}
Considering the example scenario, the contract based process of activities execution is shown in figure \ref{fig:CIM:serviceprocess}. The buy music and publish music services (update Twitter and Facebook) have pre- and post-conditions assertions that are composed into a contract for each service. The buy music pre-conditions are: (i) verify if the User data are correct; (ii) if the User is already logged in Spotify; (iii) if bank account information are correct and; (iv) if there is enough money to make the payment. As post-condition, it ensures the complete transaction and verify if a notification were sent to the user and Spotify, about the payment authorization. There are four assertions for the buy music action, and each assertion has been detailed with the assertion property and predicate that must be verified. To update services, depending of each service, there may be different restrictions. As an example, a new verification of user data and message format is appropriate (maximum 140 characters), in case of Twitter. In the case of Facebook, it is required that the user is already logged on Spotify and these data are the same as Facebook. As post-condition, it ensures that the Facebook service send a notification of success. To update Twitter a pre-condition is required, while to update Facebook is necessary a pre-condition and a confirmation notice is modelled as post-condition. As a pre-condition for ?twitter update? it is necessary that (i) the music format be correct and (ii) the twitter login and password be correct for the update.
%{\color{red} Talk about this model for the example.}
\end{example}

\begin{figure}[htpb]
\center
\includegraphics[width=0.9\textwidth]{figs/ServiceProcess.pdf}
\caption{\label{fig:CIM:serviceprocess} $\pi$-ServiceProcess model for ``To Publish Music''.}
\end{figure}
